<<<<<<< Updated upstream
---
title: "data cleaning"
author: "Alejandra Garcia Isaza"
date: '2022-09-09'
output: html_document
---

Datasets on Box folder dated 03.12.22

1MdMMedicinCUIDADORE-AnlisisSPSS03122022_DATA_NOHDRS_2022-12-04_1816.csv -> 238 columnas, 380 filas. Columnas no tienen rotulos de la encuesta, solo V1 - v238

SPSS-RawData Cuidadores 3.12.22.sav -> 243 columnas (pq?), 380 filas. Columnas no tienen rotulos de la encuesta, solo V1 - v238. Muy parecido al anterior, pero si hay algunos datos que difieren. 

Base de datos-cuidadores 03.11.22.xlsx -> 236 columnas, 72 filas. Columnas con nombre correcto.

----
1MdMMedicinCUIDADORE-Descarga010523_DATA_2023-01-05_2211.csv  --> last data download by me 01.05.23
---
Feb 10, 23

Análisis escalas-Cuidadores 03.12.22.sav --> 142 columnas, 380 filas. Columnas con nombre de las escalas.

Análisis escalaslabels-Cuidadores 03.12.22.sav --> 143 columnas, 376 filas (pq JAA quitó filas con missing data). Columnas con nombre de las escalas.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)

theme_set(theme_minimal())
```


# Loading the dataset (only analysis vars)
```{r}
d_raw <- import(here("nopublish", "1MdMMedicinCUIDADORE-1MdMMedicinCUIDADORE_DATA_2022-09-09_2130.csv"), setclass = "tbl_df")
```

#initial cleaning
```{r}

d <- d_raw[-c(1:5), ] # removing first 5 rows of data

d1 <- d %>% 
  select(-2:-34) %>% # selecting out columns with non-relevant data (, -238:-255)
  rename(c("id" = "record_id")) %>% # renaming id variable.
  clean_names() %>% # function that formats variables' names
  arrange(id) # ordering participants ids in descending order
```
# recoding missing variables as N/A
```{r include=FALSE}

# vector with missing values in dataset
missing_vals <- c(77, 99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(77, 99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(77, 99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

d2 <- recode_missing_df(d1) # the function strips out variable labels
```

# before
```{r}
view_df(d2)
```


# Adding Variable and value labels
```{r}
d3 <- d2 %>%
  mutate(sdq001_1 = as.numeric(sdq001_1),
         sdq001_1 = set_varl(sdq001_1, "Tiene en cuenta los sentimientos de otras personas"),
         sdq001_1 = set_vall(sdq001_1, c("No es cierto" = 0, "Algo cierto" = 1, "Totalmente cierto" = 2, "No responde" = 99))) # puede escribirse en ingles
```

# after
```{r}
view_df(d3)
```

# descriptive stats function
```{r include=FALSE}

# function that takes a df and returns a df with only the numeric columns
only_numeric <- function(df) {
  select_numeric <- dplyr::select_if(df, is.numeric)
  return(select_numeric)
}

# a list with the functions we want
smry <- list(n = function(x) length(x),
             n_valid = function(x) sum(!is.na(x)),
             na = function(x) sum(is.na(x)),
             M = function(x) round(mean(x, na.rm = TRUE), 2), 
             SD = function(x) round(sd(x, na.rm = TRUE), 2),
             Min = function(x) round(min(x, na.rm = TRUE), 2),
             Max = function(x) round(max(x, na.rm = TRUE), 2))

# wrapping solution in a function
descriptives <- function(df) {
  select_numeric <- only_numeric(df)
  mean_sd <- map_df(select_numeric, function(col) map_df(smry, ~.x(col)),
       .id = "column")
  return(mean_sd)
}
=======
---
title: "data cleaning"
author: "Alejandra Garcia Isaza"
date: '2022-09-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)

theme_set(theme_minimal())
```


# Loading the dataset (only analysis vars)
```{r}
d_raw <- import(here("nopublish", "1MdMMedicinCUIDADORE-1MdMMedicinCUIDADORE_DATA_2022-09-09_2130.csv"), setclass = "tbl_df")
```

#initial cleaning
```{r}

d <- d_raw[-c(1:5), ] # removing first 5 rows of data

d1 <- d %>% 
  select(-2:-34) %>% # selecting out columns with non-relevant data (, -238:-255)
  rename(c("id" = "record_id")) %>% # renaming id variable.
  clean_names() %>% # function that formats variables' names
  arrange(id) # ordering participants ids in descending order
```
# recoding missing variables as N/A
```{r include=FALSE}

# vector with missing values in dataset
missing_vals <- c(77, 99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(77, 99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(77, 99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

d2 <- recode_missing_df(d1) # the function strips out variable labels
```

# before
```{r}
view_df(d2)
```


# Adding Variable and value labels
```{r}
d3 <- d2 %>%
  mutate(sdq001_1 = as.numeric(sdq001_1),
         sdq001_1 = set_varl(sdq001_1, "Tiene en cuenta los sentimientos de otras personas"),
         sdq001_1 = set_vall(sdq001_1, c("No es cierto" = 0, "Algo cierto" = 1, "Totalmente cierto" = 2, "No responde" = 99))) # puede escribirse en ingles
```

# after
```{r}
view_df(d3)
```

# descriptive stats function
```{r include=FALSE}

# function that takes a df and returns a df with only the numeric columns
only_numeric <- function(df) {
  select_numeric <- dplyr::select_if(df, is.numeric)
  return(select_numeric)
}

# a list with the functions we want
smry <- list(n = function(x) length(x),
             n_valid = function(x) sum(!is.na(x)),
             na = function(x) sum(is.na(x)),
             M = function(x) round(mean(x, na.rm = TRUE), 2), 
             SD = function(x) round(sd(x, na.rm = TRUE), 2),
             Min = function(x) round(min(x, na.rm = TRUE), 2),
             Max = function(x) round(max(x, na.rm = TRUE), 2))

# wrapping solution in a function
descriptives <- function(df) {
  select_numeric <- only_numeric(df)
  mean_sd <- map_df(select_numeric, function(col) map_df(smry, ~.x(col)),
       .id = "column")
  return(mean_sd)
}
>>>>>>> Stashed changes
```